---
title: "Exponential Distribution in R vs Central Limit Theorem"
author: "Stefaan Delanghe"
date: "30 June 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem.We will investigate the distribution of averages of 40 exponentials. A thousand simulations will be performed.

```{r reqlib, echo=FALSE}
library(ggplot2)
set.seed(25082502)
```

## Simulations
Declaration of the simulation requirements.

```{r simulation, echo=TRUE}
nrsim <- 1000
lambda <- 0.2 
nrexp <- 40 
distmeans <- NULL 
for (i in 1 : nrsim) distmeans <- c(distmeans, mean(rexp(nrexp, lambda)))
```

## Sample Mean versus Theoretical Mean
Investigation how the sample mean is related to the theoretical mean

```{r mean, echo=TRUE}
sample.mean <- round(mean(distmeans),3);
sample.mean
theoretical.mean <- round(1/lambda,3);
theoretical.mean

hist(distmeans, main="1000 means of 40 sample exponentials", col="red", xlab = "Mean")
abline(v = sample.mean, col = "blue", lwd = 6, lty=2)
abline(v = sample.mean, col = "green", lwd = 2)
```

The histogram shows the sample (blue) and theoretical (green) mean. The sample mean is 4.989 while the theoretical mean is 5. This means that the sample mean is very close to the theoretical mean.

## Sample Variance versus Theoretical Variance
Investigation how the sample variance is compared to the theoreticam variance.

```{r variance, echo=TRUE}
theoretical.variance <- ((1/lambda)^2) / nrexp 
theoretical.variance

mean.variance <- var(distmeans) 
mean.variance
```

The theoretical variance of the distribution is 0.625 while the the sample vaiance of the distribution is 0.6081224. The difference between the theoretical and sample variance is very small.


## Distribution
Investigate if the distribution is approximately normal.

```{r distribution, echo=TRUE}
p.data <- data.frame(distmeans)
p <- ggplot(p.data, aes(x = distmeans))
p <- p + geom_histogram(aes(y=..density..), colour="black", fill="blue", bins = 30)
p <- p + labs(title="Distribution of Means", y="Density", xlab="Means")
p <- p + stat_function(fun=dnorm,args=list( mean=1/lambda, sd=sqrt(theoretical.variance)),color = "red", size = 1.0)
p <- p + stat_function(fun=dnorm,args=list( mean=mean(distmeans), sd=sqrt(mean.variance)),color = "yellow", size = 1.0)
print(p)
```

The red line is the theoretical variance as the yellow line is the mean variance. The two lines are pretty close aligned with each other. This close alignement indicates that the distribution is approximately normal.
